---
title: "data-exploration"
author: "Alissa Ji"
date: "7/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load relevant libraries:
```{r}
library(survey)
library(tidyverse)
```

Load data:
```{r}
# housing <- read.csv("ahs2017_hhsubset_r.csv")
housing <- read.csv("ahs2017_flat_r.csv")
```

# Weighted exploratory analysis

Adding weights:
```{r}
housing.w <- svydesign(ids = ~1, data = housing, weights = housing$WEIGHT)
# summary(housing.w)
```

```{r}
# HHSEX
prop.table(table(housing$HHSEX))
prop.table(svytable(~HHSEX, design = housing.w))
# prop.table(svytable(~HHSEX + HHRACE, design = housing.w), 2)
```

# Cross-checking against Census FactFinder

## Household race definition #1: All persons in household identify with a given race
Sample size in each region:
```{r}
housing %>% group_by(HOUSEHOLDRACE) %>% summarize(n.count = n())
```
## Household race definition #2: Only householder identifies with a given race
Sample size in each region:
```{r}
housing %>% group_by(HHRACE3) %>% summarize(n.count = n())
```

## Racial breakdown/eth:
```{r}
# by #1: entire houshold
prop.house.race <- prop.table(svytable(~HOUSEHOLDRACE, design = housing.w)) # excluding Hispanic Whites
prop.house.raceeth <- prop.table(svytable(~HOUSEHOLDRACEETH, design = housing.w)) # excluding Hispanic Whites
prop.house.span <- prop.table(svytable(~HOUSEHOLDSPAN, design = housing.w))

# by #2: householder
prop.hh.race <- prop.table(svytable(~HHRACE3, design = housing.w)) # excluding Hispanic Whites
prop.hh.raceeth <- prop.table(svytable(~RACEETH, design = housing.w)) # excluding Hispanic Whites
prop.hh.span <- prop.table(svytable(~HHSPAN2, design = housing.w))

# from Fact Finder
race.ff <- c(2726278, 18215328, 41393491, 608219, 235507457)
names(race.ff) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                    "White Alone")
prop.race.ff <- race.ff / sum(race.ff)

raceeth.ff <- c(2726278, 18215328, 41393491, 608219, 197285202)
names(raceeth.ff) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                    "White NH Alone")
prop.raceeth.ff <- raceeth.ff / sum(raceeth.ff)

hisp.ff <- 56510571
total.pop.ff <- 321004407
prop.hisp.ff <- hisp.ff / total.pop.ff

# Calculate % difference
## def #1:
(prop.house.race - prop.race.ff) 
(prop.house.raceeth - prop.raceeth.ff)
(prop.house.span[1] - prop.hisp.ff)

## def #2
(prop.hh.race - prop.race.ff)
(prop.hh.raceeth - prop.raceeth.ff)
(prop.hh.span[1] - prop.hisp.ff)
```

Household race (def #1) numbers have margin of error > 25% for:
* Asian households: -0.24888033
* NHPI households: 0.34009632

Householder race (def #2) numbers have much smaller margin of error for white and Black population estimates but larger margin of error for the following groups, >25%:
* AIAN: 0.356737430
* Asian: -0.193478359
* NHPI: 0.786307222

Unsurprising that race at level of household doesn't entire match 2017 ACS data because AHS conducted at level of the household.

## HH income by race/eth:
** Note: running into issues here **
```{r}
# by #1: household
inc.med.byhouserace <- svyby(~HINCP, ~HOUSEHOLDRACE, housing.w, svyquantile, quantiles=0.5, na.rm = FALSE)
inc.med.byhousespan <- svyby(~HINCP, ~HOUSEHOLDSPAN, housing.w, svymean)

# by #2: householder
inc.med.byhhrace <- svyby(~HINCP, ~HHRACE3, housing.w, svyquantile, quantiles=0.5, na.rm = FALSE)
inc.med.byhhspan <- svyby(~HINCP, ~HHSPAN2, housing.w, svyquantile, quantiles=0.5, na.rm = FALSE)

# from Fact Finder (inflation-adjusted dollars)
inc.med.ff <- c(41882, 83456, 40232, 60734, 63704)
names(inc.med.ff) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                       "White Alone")
inc.span.rd <- 49793

# Calculate % difference
(inc.med.byhouserace - inc.med.ff) / inc.med.ff
(inc.med.byhhrace - inc.med.ff) / inc.med.ff

(inc.med.housingspan[1] - inc.span.ff) / inc.span.ff
(inc.med.hhspan[1] - inc.span.ff) / inc.span.ff
```

## Homeownership by race/eth:
```{r}
# by #1: entire houshold
tenure.byhousingrace <- prop.table(svytable(~HOUSEHOLDRACE + TENURE, design = housing.w), 1)
tenure.byhousingspan <- prop.table(svytable(~HOUSEHOLDSPAN + TENURE, design = housing.w), 1)


# by #2: householder
tenure.byhhrace <- prop.table(svytable(~HHRACE3 + TENURE, design = housing.w), 1)
tenure.byhhspan <- prop.table(svytable(~HHSPAN2 + TENURE, design = housing.w), 1)

# from Fact Finder
tenure.ff <- matrix(c(0.544, 0.456, 
                      0.596, 0.404, 
                      0.414, 0.586, 
                      0.412, 0.588, 
                      0.695, 0.305),
                    nrow = 5, ncol = 2, byrow = TRUE)
rownames(tenure.ff) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                       "White Alone")
colnames(tenure.ff) <- c("Owners", "Renters")

tenure.span.ff <- c(0.472, 0.528)
names(tenure.span.ff) <- c("Owners", "Renters")

# Calculate % difference
## by #1:
(tenure.byhousingrace[,2:3] - tenure.ff)
(tenure.byhousingspan[1,2:3] - tenure.span.ff)

## by #2:
(tenure.byhhrace[,2:3] - tenure.ff)
(tenure.byhhspan[1,2:3] - tenure.span.ff)
```

* Similar margin of error across race definitions, but def #2 has slightly smaller error for AIAN and NHPI households, likely due to greater sample size
* Margin of error reasonable, nearing 25% for AIAN households in def #1

# Additional exploratory analysis
Geographic divisions:
```{r}
# Geographic division by race
prop.table(table(housing$HHRACE2, housing$DIVISION), 2)

# Geographic division by race and ethnicity
prop.table(table(housing$RACEETH, housing$DIVISION), 2)

# Geographic division by Asian ethnicity
prop.table(table(housing$HHRACEAS, housing$DIVISION), 2) # of those who identify as Asian
```

Types of housing and federal assistance (comparing different measures of housing assistance):
```{r}
table(housing$TENURE)
table(housing$BLD2) # mostly single family homes

# of those who are renters (NOT owners)
# forms of federal assistance and rent control
table(housing$RENTCNTRL, housing$HUDSUB) # some individuals who have federal assistance are not under rent control or stabilization
# this indicates that a large number who are eligible did not report or do not receive assistance

table(housing$RENTSUB, housing$HUDSUB)
```

*Idea*: demonstrate how many people are eligible for federal housing assistance vs. receive assistance (RENTSUB), by race and geography

General housing:
```{r}
# by race
prop.table(table(housing$RACEETH, housing$TENURE), 1)
prop.table(table(housing$HHRACEAS, housing$TENURE), 1)

# by geography
prop.table(table(housing$DIVISION, housing$TENURE), 1)
```

Federal housing assistance (renters):
```{r}
# by race and ethnicity
table(housing$RACEETH, housing$HUDSUB)
# prop.table(table(housing$RACEETH, housing$HUDSUB), 1)
prop.table(table(housing$RACEETH, housing$HUDSUB), 2)
## Asian only: much lower rate of living in public housing / voucher recipients

# by Asian ethnic group
table(housing$HHRACEAS, housing$HUDSUB)
# prop.table(table(housing$HHRACEAS, housing$HUDSUB), 1)
prop.table(table(housing$HHRACEAS, housing$HUDSUB), 2)

# by geography
table(housing$HUDSUB, housing$DIVISION)
prop.table(table(housing$HUDSUB, housing$DIVISION), 2)

# three-way crosstab: federal housing assistance AND Asian ethnic group by geography
levels(housing$HUDSUB)[2] <- "Public or private"
t1 <- xtabs(~DIVISION+RACEETH+HUDSUB, data=housing)
ftable(t1)
```

Monthly housing costs
```{r}
# by race
tapply(housing$TOTHCAMT, housing$HHRACE2, mean)

# by Asian ethnic group
tapply(housing$TOTHCAMT, housing$HHRACEAS, mean)
```

*Idea*: could calculate monthly housing costs as proportion of income
*Idea*: renters vs. owners

# Sample sizes

Sample size in each region:
```{r}
ahs_subset %>% group_by(DIVISION) %>% summarize(n.count = n())
```
Regions with highest number of households sampled: 
(1) South Atlantic, n=14257;
(2) Pacific, n=13330;
(3) East North Central, n=9408

Regions with smallest number of households sampled:
(1) East South Central, n=2752;
(2) West North Central, n=3036;
(3) New England, n=3577

Sample size by race and ethnicity:
```{r}
ahs_subset %>% group_by(RACEETH) %>% summarize(n.count = n())

ahs_subset %>% group_by(HHSPAN2) %>% summarize(n.count = n())

```

# Cross-checking against Racial Data

## Demographics of Head of Household

Racial breakdown:
```{r}
prop.raceeth <- prop.table(table(housing$RACEETH)) # excluding Hispanic Whites
prop.hhspan2 <- prop.table(table(housing$HHSPAN2))

# from Racial Data
total.pop.rd <- 321004407
race.rd <- c(2632102, 17186320, 40610815, 570116, 197277789)
race.rd <- race.rd / sum(race.rd)
names(race.rd) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                    "White Alone NH")
hisp.rd <- 56510571
prop.hisp.rd <- hisp.rd / total.pop.rd

# Calculate % difference
(prop.raceeth - race.rd) / race.rd
prop.hhspan2[1] - prop.hisp.rd
```

Margin of error > 25%:
AIAN Alone error: 35.667%
NHPI Alone error: 122.371%

HH income by race:
```{r}
inc.mean.byrace <- tapply(housing$HINCP, housing$RACEETH, mean) # mean
inc.med.byrace <- tapply(housing$HINCP, housing$RACEETH, median) # median

inc.mean.hhspan <- tapply(housing$HINCP, housing$HHSPAN2, mean) # mean
inc.med.hhspan <- tapply(housing$HINCP, housing$HHSPAN2, median) # median

# from Racial Data
inc.med.rd <- c(40315, 80398, 38183, 57372, 63256)
names(inc.med.rd) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                       "White Alone NH")
inc.span.rd <- 46627

# Calculate % difference
(inc.med.byrace - inc.med.rd) / inc.med.rd
(inc.med.hhspan[1] - inc.span.rd) / inc.span2.rd
```

All incomes within 25% of Racial Data estimates. The highest margin of error is -12.265% for Black Alone, indicating that the AHS survey results underestimated household incomes as compared to the ACS.

Age by race:
```{r}
age.byrace <- tapply(housing$HHAGE, housing$RACEETH, median)
age.byspan <- tapply(housing$HHAGE, housing$HHSPAN2, median)

# from Racial Data
age.rd <- c(32.5, 36.7, 33.6, 31.4, 43.2)
names(age.rd) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                       "White Alone NH")
age.span.rd <- 28.7

# Calculate % difference
(age.byrace - age.rd) / age.rd
(age.byspan[1] - age.span.rd) / age.span.rd
```

Margin of error on ages of householder are all > 25%, which is expected because the subpopulation of householders is likely much older than the overall population.

Homeownership by race:
```{r}
tenure.byrace <- prop.table(table(housing$RACEETH, housing$TENURE), 1)
tenure.byspan <- prop.table(table(housing$HHSPAN2, housing$TENURE), 1)

# from Racial Data
tenure.rd <- matrix(c(0.541, 0.587, 0.419, 0.401, 0.716, 
                    0.459, 0.413, 0.581, 0.599, 0.284),
                    nrow = 5, ncol = 2, byrow = TRUE)
rownames(tenure.rd) <- c("AIAN Alone", "Asian Alone", "Black Alone", "NHPI Alone",
                       "White Alone NH")
colnames(tenure.rd) <- c("Owners", "Renters")

tenure.span.rd <- c(0.464, 0.536)
names(tenure.span.rd) <- c("Owners", "Renters")

# Calculate % difference
(tenure.byrace[,2:3] - tenure.rd) / tenure.rd
(tenure.byspan[1,2:3] - tenure.span.rd) / tenure.span.rd
```

AHS margin of error > 25% for:
* Asian Alone, Owner
* Black Alone, Owner and Renter (most significant)

## Demographics of Household

